---
title: "Course 5 Project 1"
author: "SK Sharma"
date: "September 28, 2018"
output: html_document
---
# Project 1

## Code for reading in the dataset and/or processing the data.

```{r}
if(!dir.exists("c5p1")){dir.create("c5p1")}
 setwd("./c5p1")
url<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
 if(!file.exists("c5p1.zip")){download.file(url, destfile = "c5p1.zip")}
 if(!file.exists("activity.csv")){unzip("c5p1.zip")}
data<-read.csv("activity.csv", sep = ",", header = TRUE)
head(data)
```

## Histogram of the total number of steps taken each day.

```{r}
dsteps<-tapply(data$steps, data$date, sum, na.rm=TRUE)
head(dsteps)
hist(dsteps, main = "Histogram of the total number of steps taken each day", xlab = "Numbers of steps taken")
```

## Mean and median number of steps taken each day.

```{r}
dsteps<-data.frame(dsteps)
colnames(dsteps)[1]<-"Dates"
mean(dsteps$Dates)
median(dsteps$Dates)
```


## Time series plot of the average number of steps taken.

```{r}
dinterval<-tapply(data$steps, data$interval, mean, na.rm=TRUE)
dinterval<-data.frame(dinterval)
dataint<-data.frame(interval=row.names(dinterval), steps=dinterval, row.names = NULL)
plot(as.numeric(as.character(dataint$interval)), dataint$dinterval, type="l", xlab="Interval", ylab="Mean Steps")
```

## The 5-minute interval that, on average, contains the maximum number of steps


```{r}
dataint[which.max(dataint$dinterval),]
```



## 6.  The total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)


```{r}
sum(!complete.cases(data$steps))
```

## Code to describe and show a strategy for imputing missing data

```{r}
dataf<-data
for(i in 1 : length(dataf$steps)){
        if(is.na(dataf[i,1])){
        r<-which(dataint$interval==dataf[i,3])
        dataf[i,1]<-dataint[r,2]}
        }
head(dataf)
```


## Histogram of the total number of steps taken each day and  the mean and median total number of steps taken per day.
        

```{r}
dfsteps<-data.frame(steps=tapply(dataf$steps, dataf$date, sum),date=levels(factor(dataf$date)),row.names = NULL)
head(dfsteps)
hist(dfsteps$steps, main = "Histogram of the total number of steps taken each day (Imputed Data)", xlab = "Numbers of steps taken")
mean(dfsteps$steps)
median(dfsteps$steps)
```

- Do these values differ from the estimates from the first part of the assignment?

Data slightly differ one can observe this by comparing the means and meadians which increases in the imputed data.

- What is the impact of imputing missing data on the estimates of the total daily number of steps?

Impact is that the frequency in the imputed data in the 1 and 3 interval decrease. (one can observe this by compaing two histograms.)




## Creating a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
dataw<-data.frame(cbind(dataf,Days=weekdays(as.Date(dataf$date))))
dataw$Days<-ifelse(dataw$Days=="Saturday"|dataw$Days=="Sunday", "weekend","weekday")
head(dataw)
levels(as.factor(dataw$Days))
```


## Making a panel plot containing a time series plot  of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
library(dplyr)
datap<-transform(dataw, Days=factor(Days))
datafp<-aggregate(steps ~ interval + Days, datap, mean)
library(lattice)
xyplot(datafp$steps~datafp$interval|datafp$Days, layout=c(1,2), type="l", xlab = "Intervals", ylab = "Numbers of Steps")
```
